Dataflow-Guided Filtering for Efficient and Adjustable Run-Time Monitoring

### 2014-05-31 arch.tex
Various typos/fixes.

For both filtering and partial monitoring, we need to be able to track
dataflows. For filtering, we need to track the dataflow of null metadata and
for partial monitoring, we need to track the dataflow of invalid metadata.
In this section, we will describe our dataflow-guided hardware architecture for
enabling both filtering and partial monitoring.
Figure~\ref{fig:arch.overview} shows a high-level block diagram of the
architecture. The basic idea is to insert this dataflow-guided hardware between the main core and
the monitoring core that will reduce the number of monitoring events that reach
the monitoring core and thus reduce the overheads of monitoring.

@@

% Overview
In order to perform filtering, we need to be able to identify and track flows
of null metadata. Similarly, for partial monitoring,
 we need to be able to track the flow of invalid metadata. In
order to efficiently these flows, we use a dedicated hardware
dataflow engine. A block diagram of the dataflow engine is shown in
Figure~\ref{fig:arch.dataflow}.

% Flags and flag storage
In order track both null metadata flows and invalid metadata flows, the
dataflow engine tracks the flow of 2-bit flags. One bit is used to
mark metadata as null while the other bit is used to mark metadata as invalid.
Monitoring schemes typically keep metadata corresponding to the main core's
register file as well as the main core's memory space. Similarly, the dataflow
engine uses an on-chip register file to store flags for register file metadata.
In addition, a set of flags are stored that correspond to memory metadata.
These memory metadata flags are accessed through a cache and backed to main
memory. All flags are initialized as null and valid when the system starts.

@@

Metadata is marked as non-null when the monitoring core writes a valid metadata
value. When the monitoring core performs these metadata update operations, it
also writes to the corresponding dataflow flag to mark the flag as non-null.

@@

In terms of filtering using null flags, array allocations are never filtered
since these will always involve non-null metadata (i.e., setting new base and
bounds). For ALU operations, if both source operands are null, then the event
can be filtered by setting the destination register as null. For load and
store, only a single source needs to be checked to be null in order for the
event to be filtered. Similarly, the destination register or memory location
needs to be marked as null if the event is filtered.

@@

Another limitation is that the dataflow engine, as described, only marks one
flag location on each monitoring event. Similarly, the dataflow engine can only
read in two sets of flags in order to determine whether an event can be
filtered or not. This was done because we found that for many monitoring
schemes, the monitoring operations corresponded to program instructions, which
read up to two operands and write to one destination. For monitoring schemes
which update multiple pieces of metadata on a monitoring event, it may be
adequate to associate a null and invalid flag with a single piece of metadata.  For
example, a scheme may use a data structure that includes several pieces of
related metadata. On a drop, one of these pieces could be marked as invalid in
order to indicate that the entire structure is invalid.  Similarly, marking one
specified piece as null may indicate that the entire data structure's metadata
is null.  Alternatively, the read and write capabilities of the dataflow
engine could be expanded to support these more complex monitoring schemes.

>>>>

For both filtering and partial monitoring, we need to be able to track
dataflows. For filtering, we need to track the dataflow of null metadata and
for partial monitoring, we need to track the dataflow of invalid metadata.
In this section, we will describe our dataflow-guided hardware architecture for
enabling both filtering and partial monitoring.
Figure~\ref{fig:arch.overview} shows a high-level block diagram of the
architecture. The basic idea is to insert this dataflow-guided hardware between the main core and
the monitoring core. This hardware reduces the number of monitoring events that reach
the monitoring core in order to reduce monitoring overheads.

@@

% Overview
In order to perform filtering, we need to be able to identify and track flows
of null metadata. Similarly, for partial monitoring,
 we need to be able to track the flow of invalid metadata. In
order to efficiently track these flows, we use a dedicated hardware
dataflow engine. A block diagram of the dataflow engine is shown in
Figure~\ref{fig:arch.dataflow}.

% Flags and flag storage
In order to track both null metadata flows and invalid metadata flows, the
dataflow engine tracks the flow of 2-bit flags. One bit is used to
mark metadata as null while the other bit is used to mark metadata as invalid.
Monitoring schemes typically keep metadata corresponding to the main core's
register file as well as the main core's memory space. Similarly, the dataflow
engine uses an on-chip register file to store flags for register file metadata.
In addition, a set of flags are stored that correspond to memory metadata.
These memory metadata flags are accessed through a cache and backed to main
memory. All flags are initialized as null and valid when the system starts.

@@

Metadata is marked as non-null when the monitoring core writes a valid metadata
value. When the monitoring core performs these metadata update operations, it
also writes to the corresponding dataflow flag to mark the flag as non-null. In
addition, it also re-validates the flag if it was set as invalid due to a
dropped monitoring event.

@@

In terms of filtering using null flags, array allocations are never filtered
since these will always involve non-null metadata (i.e., setting new base and
bounds). For ALU operations, if both source operands are null, then the event
can be filtered by setting the destination register as null. For loads and
stores, only a single source needs to be checked to be null in order for the
event to be filtered. Similarly, the destination register or memory location
needs to be marked as null if the event is filtered.

@@

Another limitation is that the dataflow engine, as described, only marks one
flag location on each monitoring event. Similarly, the dataflow engine can only
read in two sets of flags in order to determine whether an event can be
filtered or not. This was done because we found that for many monitoring
schemes, the monitoring operations corresponded to program instructions, which
read up to two operands and write to one destination. For monitoring schemes
which update multiple pieces of metadata on a monitoring event, it may be
adequate to associate a null and invalid flag with a single piece of metadata.  For
example, a scheme may use a data structure that includes several pieces of
related metadata. On a drop, one portion of the data structure could be marked as invalid in
order to indicate that the entire structure is invalid.  Similarly, marking one
specified piece as null may indicate that the entire data structure's metadata
is null.  Alternatively, the read and write capabilities of the dataflow
engine could be expanded to support these more complex monitoring schemes.

### 2014-05-31 evaluation.tex
Various typos/fixes.

We implemented dataflow-guided monitoring architecture by
modifying the ARM version of the gem5 simulator \cite{gem5} to support parallel
run-time monitoring. We model the main and monitoring cores as running at 2.5
GHz with 4-way set-associative private L1 I/D caches and a shared 8-way 2 MB L2
cache. This setup is similar to the Snapdragon 801 processor commonly found in
mobile systems. The dataflow engine uses a 1 kB cache for flags.

In order to explore the generality of the architecture for
different monitors, we implemented three different monitors: uninitialized
memory check (UMC), array bounds check (BC), and dynamic information flow
tracking (DIFT).  Uninitialized memory check seeks to detect loading from
memory locations that are not initialized first.  Array bounds check, as
mentioned in Section~\ref{sec:monitoring} is a monitoring scheme that aims to
detect buffer overflows where memory accesses go beyond the boundaries of an
array. We modify the implementation of {\tt malloc} to set base and bound
metadata information. Dynamic information flow tracking is a security
monitoring scheme,
which detects when information from untrusted sources is used to affect the
program control flow. For the benchmarks we consider, we mark data read from
files as untrusted. We implemented a multi-bit DIFT scheme which marks
untrusted data with a 32-bit metadata identifier so
that if an error is detected, it is possible to have information about where
the data originated from.

We tested our system using all C benchmarks from SPECint
CPU2006 \cite{spec2006}. We focus on the C benchmarks because of our
modification of {\tt malloc} to support array bounds check. Although we do not
show results for the C++ benchmarks, we note that the results for UMC and DIFT
for these benchmarks are similar to the other results shown. For each
benchmark, we simulated for 1 billion instructions. An initial slack of 1
million cycles, which is less than 1\% of
the total execution time, is given for a 10\% target overhead. This initial
slack is scaled proportionally for other overhead targets.

\subsection{Baseline Monitoring Overheads}
% Full monitoring overheads
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_full_mon.pdf}
    \vspace{-0.2in}
    \caption{Full monitoring overheads for UMC, BC, and DIFT.}
    \label{fig:evaluation.full_mon}
    \vspace{-0.1in}
  \end{center}
\end{figure} Figure~\ref{fig:evaluation.full_mon} shows the execution times of
performing full monitoring normalized to the execution times of the benchmarks
without monitoring. In these results, no filtering or partial monitoring is
done. UMC shows normalized execution times from 4x to 11x with an average
of 7x. BC shows normalized execution times of 9-28x with an average of
18x while DIFT shows normalized execution times of 6x-17x with an average of
11x. One of the reasons for these high overheads is that our implementations of
these monitors dynamically allocate memory for new metadata. This can be an
expensive, many-cycle operation. Statically allocating memory for metadata
will reduce the execution time but requires a large upfront memory footprint.

@@

Figure~\ref{fig:evaluation.filtering} shows the
normalized execution times with filtering enabled. We see significant
reductions in overhead for all three monitoring schemes. UMC sees normalized
execution times 2-6x with an average of 4x with null metadata filtering. For BC, normalized execution
times drop from 18x down to 4x on average. The range of normalized execution
times for filtered BC is 1.1-17x. This is not surprising since in the baseline
implementation all loads and stores needed to be monitored. However, with
filtering, only loads and stores corresponding to arrays need to be forwarded.
Finally, DIFT sees the largest reduction in overheads with only 13\% overheads
on average after filtering with a maximum of 42\%. This is due, in part, to the fact that
for our implementation of DIFT on SPEC
benchmarks, we only mark data read from files as tainted. For most of these
benchmarks, this propagates to relatively few instructions. Instead, if we
targeted network or streaming applications, which have larger amounts in
untrusted input data, we would expect to see less filtering.

@@

Although the overheads of DIFT are quite low after filtering, BC and UMC still
show significant overheads. In this section, we evaluate the effectiveness of
using partial monitoring (in addition to filtering) to trade-off coverage for further reduced overheads.
Figure~\ref{fig:evaluation.bc_sweep} shows the monitoring coverage achieved by
array bounds check as we vary the overhead budget. We measure monitoring
coverage monitoring check operations that are still performed as a percentage
of the total number of monitoring check operations. This number does not
include the number metadata update operations that are done and serves as a
better metric for the actual coverage of the security or reliability technique.
The overhead is measured as a percentage of total number of checks performed
with full monitoring. The results shown here are using an unrestricted dropping
policy.

@@

Figure~\ref{fig:evaluation.umc_sweep} shows the analogous graph for UMC. Again
we see that varying overhead budgets enables partial monitoring. We see that
with a 10\% overhead budget, UMC achieves 32\% monitoring coverage on average
and with a 50\% budget, UMC achieves 51\% coverage.

@@

% UMC exec time
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_umc_exec_time.pdf}
    \vspace{-0.2in}
    \caption{Error in meeting overhead budget for uninitialized memory check.}
    \label{fig:evaluation.umc_exec_time}
    \vspace{-0.2in}
  \end{center}
\end{figure}

% UMC coverage across policies
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_umc_coverage.pdf}
    \vspace{-0.2in}
    \caption{Coverage for uninitialized memory check for different dropping policies.}
    \label{fig:evaluation.umc_coverage}
    \vspace{-0.1in}
  \end{center}
\end{figure}
In this section we evaluate the trade-offs of unrestricted dropping and source-only dropping.
Figure~\ref{fig:evaluation.umc_exec_time} shows the difference between the
overhead budget and the run-time monitoring overheads for UMC. A positive value
means that the overhead target was overshot while a negative value indicates
that the overhead budget was met. We see similar results for unrestricted
dropping and source-only dropping due to the fact that UMC consists of a large
number of independent monitoring flows.
Figure~\ref{fig:evaluation.umc_coverage} shows the coverage of UMC for
unrestricted dropping and source-only dropping. We see that source-dropping
consistently achieves higher coverage than unrestricted dropping. This is due
to the fact that some of the overheads of monitoring for unrestricted dropping
are being spent on wasted work as discussed in Section~\ref{sec:policies.events}.

% BC exec time
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_bc_exec_time.pdf}
    \vspace{-0.2in}
    \caption{Error in meeting overhead budget for array bounds check.}
    \label{fig:evaluation.bc_exec_time}
    \vspace{-0.2in}
  \end{center}
\end{figure}

@@

% Multi-run coverage with unrestricted dropping
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_multirun_coverage.pdf}
    \vspace{-0.2in}
    \caption{Coverage over multiple runs for BC at 20\% overhead.}
    \label{fig:evaluation.multirun}
    \vspace{-0.1in}
  \end{center}
\end{figure}

One possible application of partial monitoring with low overheads is to enable
cooperative debugging. The idea with cooperative debugging is to enable partial
monitoring with very low overheads across a large number of users or runs. By
varying the pattern of partial monitoring done on each run, the goal is to
achieve high or total coverage across all runs. Varying the monitoring that is
done for different runs can be achieved by introducing some randomness into the
dropping decision. Figure~\ref{fig:evaluation.multirun} shows the total
coverage achieved over multiple runs for array bounds check running with a 20\%
overhead budget and an unrestricted dropping policy. Here, when events are
dropped probabilistically with the probability scaled proportionally to slack.
That is, when slack is high, the probability of dropping is low and the
probability increases as we run out of slack. We see that the coverage achieved
quickly asymptotes for an unrestricted dropping policy. This is due to the fact
that since there are multiple dependent monitoring operations that must be done
in order for a final check to be successful, a small amount of randomness at
each event quickly compounds to reduce the probability of certain checks to
near zero. Instead, source-only dropping is much better suited for achieve high
coverage over multiple runs.

@@

Adding the dataflow engine in order to enable filtering and adjustable adjustable adds
overheads in terms of area and power. We use McPat \cite{mcpat-micro09} to get
a first-order estimate of these area and power overheads in a 40 nm technology
node. McPat estimates the main core area as 2.71 mm$^2$ and the peak power usage as
965 mW averaged across all benchmarks. The average runtime power usage was 385 
mW. These area and power numbers consist of the core and
L1 cache, but do not include L2 cache, memory controller, and other
peripherals. The power numbers include dynamic as well as static (leakage)
power. For the dataflow engine, we modeled the ALUs used for address
calculation, the dataflow flag register file and cache, the configuration
tables, and the filter decision table. These were modeled using the
corresponding memory and ALU objects in McPat. We
note that this is only a rough area and power result since components such as the
wires connecting these modules have not been modeled. However, this gives a
sense of the order-of-magnitude overheads involved with implementing our
approach.

Our results show that an additional 0.197 mm$^2$ of silicon area is needed, an
increase of 7\% of the main core area. Table~\ref{tab:evaluation.area_power}
shows the peak and runtime power overheads averaged across all benchmarks with
a 50\% monitoring overhead target. The peak power is 5-7 mW, which is
5-7\% of the main core's peak power usage. The average runtime power is 3-4 
mW, corresponding to 8-11\% of the main core's runtime power.

>>>>

We implemented our dataflow-guided monitoring architecture by
modifying the ARM version of the gem5 simulator \cite{gem5} to support parallel
run-time monitoring. We model the main and monitoring cores as running at 2.5
GHz with 4-way set-associative private L1 I/D caches and a shared 8-way 2 MB L2
cache. This setup is similar to the Snapdragon 801 processor commonly found in
mobile systems. The dataflow engine uses a 1 kB cache for flags.

In order to explore the generality of the architecture for
different monitors, we implemented three different monitors: uninitialized
memory check (UMC), array bounds check (BC), and dynamic information flow
tracking (DIFT).  Uninitialized memory check seeks to detect loading from
memory locations that are not initialized first.  Array bounds check, as
mentioned in Section~\ref{sec:monitoring}, is a monitoring scheme that aims to
detect buffer overflows where memory accesses go beyond the boundaries of an
array. We modify the implementation of {\tt malloc} to set base and bound
metadata information. Dynamic information flow tracking is a security
monitoring scheme
which detects when information from untrusted sources is used to affect the
program control flow (i.e., indirect control instructions). For the benchmarks we consider, we mark data read from
files as untrusted. We implemented a multi-bit DIFT scheme which marks
untrusted data with a 32-bit metadata identifier so
that if an error is detected, it is possible to have information about where
the data originated from.

We tested our system using all C benchmarks from SPECint
CPU2006 \cite{spec2006}. Since our implementation of BC depends on the
modification of {\tt malloc} to set array bounds information, we focus on the C
SPECint benchmarks. Although we do not
show results for the C++ benchmarks, we note that the results for UMC and DIFT
for these benchmarks are similar to the other results shown. For each
benchmark, we simulated for 1 billion instructions. An initial slack of 1
million cycles, which is less than 1\% of
the total execution time, is given for a 10\% target overhead. This initial
slack is scaled proportionally for other overhead targets.

\subsection{Baseline Monitoring Overheads}
% Full monitoring overheads
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_full_mon.pdf}
    \vspace{-0.2in}
    \caption{Full monitoring overheads for UMC, BC, and DIFT.}
    \label{fig:evaluation.full_mon}
    \vspace{-0.1in}
  \end{center}
\end{figure} Figure~\ref{fig:evaluation.full_mon} shows the execution times of
performing full monitoring normalized to the execution times of the benchmarks
without monitoring. In these results, no filtering or partial monitoring is
done. UMC shows normalized execution times from 4x to 11x with an average
of 7x. BC shows normalized execution times of 9-28x with an average of
18x while DIFT shows normalized execution times of 6x-17x with an average of
11x. One of the reasons for these high overheads is that our implementations of
these monitors dynamically allocate memory for new metadata. This can be an
expensive, many-cycle operation. Statically allocating memory for metadata
will reduce the execution time but requires a large upfront memory footprint
which caused issues with our simulator.

@@

Figure~\ref{fig:evaluation.filtering} shows the
normalized execution times with filtering enabled. We see significant
reductions in overhead for all three monitoring schemes. UMC sees normalized
execution times of 2-6x with an average of 4x with null metadata filtering. For BC, normalized execution
times drop from 18x down to 4x on average. The range of normalized execution
times for filtered BC is 1.1-17x. This is not surprising since in the baseline
implementation all loads and stores needed to be monitored. However, with
filtering, only loads and stores corresponding to arrays need to be forwarded.
Finally, DIFT sees the largest reduction in overheads with only 13\% overheads
on average after filtering, with a maximum of 42\%. This is due, in part, to the fact that
for our implementation of DIFT on SPEC
benchmarks, we only mark data read from files as tainted. For most of these
benchmarks, this propagates to relatively few instructions. Instead, if we
targeted network or streaming applications, which have larger amounts of
untrusted input data, we would expect to see less filtering.

@@

Although the overheads of DIFT are quite low after filtering, BC and UMC still
show significant overheads. In this section, we evaluate the effectiveness of
using partial monitoring (in addition to filtering) to trade-off coverage for further reduced overheads.
Figure~\ref{fig:evaluation.bc_sweep} shows the monitoring coverage achieved by
array bounds check as we vary the overhead budget. 
We define \emph{monitoring coverage} as the 
percentage of checks that are performed 
(indirect jumps in DIFT, loads in UMC, and memory accesses in BC). 
The metric is chosen to understand
how likely an error/attack instance is to be detected on an individual system. 
While we could not evaluate real errors/attacks due to difficulty in setting up
a large number of bugs and exploits, we believe that the percentage of checks
provides a good estimate of detection probability when errors/attacks are 
uniformly distributed across checks.
Note that this metric is different from the
percentage of monitoring events that are not dropped, which includes
non-check instructions.

@@

Figure~\ref{fig:evaluation.umc_sweep} shows the analogous graph for UMC. Again
we see that varying overhead budgets enables partial monitoring. We see that
with a 10\% overhead budget, UMC achieves 32\% monitoring coverage on average.
By increasing this overhead budget to 50\%, UMC achieves 51\% coverage. Even
higher coverage can be achieved by allowing higher overheads.

@@

% UMC exec time
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_umc_exec_time.pdf}
    \vspace{-0.2in}
    \caption{Error in meeting overhead budget for uninitialized memory check for different dropping policies.}
    \label{fig:evaluation.umc_exec_time}
    \vspace{-0.2in}
  \end{center}
\end{figure}

% UMC coverage across policies
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_umc_coverage.pdf}
    \vspace{-0.2in}
    \caption{Coverage for uninitialized memory check for different dropping policies.}
    \label{fig:evaluation.umc_coverage}
    \vspace{-0.1in}
  \end{center}
\end{figure}
In this section we evaluate the trade-offs between unrestricted dropping and source-only dropping.
Figure~\ref{fig:evaluation.umc_exec_time} shows the difference between the
overhead budget and the run-time monitoring overheads for UMC. A positive value
means that the overhead target was overshot while a negative value indicates
that the overhead budget was met. We see similar results for unrestricted
dropping and source-only dropping due to the fact that UMC consists of a large
number of independent monitoring flows.
Figure~\ref{fig:evaluation.umc_coverage} shows the coverage of UMC for
unrestricted dropping and source-only dropping. We see that source-dropping
consistently achieves higher coverage than unrestricted dropping. This is due
to the fact that some of the overheads of monitoring for unrestricted dropping
are being spent on wasted work as discussed in Section~\ref{sec:policies.events}.

% BC exec time
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_bc_exec_time.pdf}
    \vspace{-0.2in}
    \caption{Error in meeting overhead budget for array bounds check for different dropping policies.}
    \label{fig:evaluation.bc_exec_time}
    \vspace{-0.2in}
  \end{center}
\end{figure}

@@

% Multi-run coverage with unrestricted dropping
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_multirun_coverage.pdf}
    \vspace{-0.2in}
    \caption{Coverage over multiple runs for BC at 10\% overhead.}
    \label{fig:evaluation.multirun}
    \vspace{-0.1in}
  \end{center}
\end{figure}

One possible application of partial monitoring with low overheads is to enable
cooperative debugging. The idea with cooperative debugging is to use partial
monitoring with very low overheads across a large number of users or runs. By
varying the pattern of partial monitoring done on each run, the goal is to
achieve high coverage across multiple runs. Varying the monitoring that is
done for different runs can be achieved by introducing some randomness into the
dropping decision. Figure~\ref{fig:evaluation.multirun} shows the total
coverage achieved over multiple runs for array bounds check running with a 10\%
overhead budget and an unrestricted dropping policy. Each run was simulated for
100 million instructions. For these simulations, events were
dropped probabilistically with the probability scaled proportionally to slack.
That is, when slack is high, the probability of dropping is low and the
probability increases as we run out of slack. We see that the coverage achieved
quickly asymptotes for an unrestricted dropping policy. This is due to the fact
that since there are multiple dependent monitoring operations that must be done
in order for a final check to be successful, a small amount of randomness at
each event quickly compounds to reduce the probability of certain checks to
near zero. Instead, source-only dropping is much better suited for achieving high
coverage over multiple runs.

@@

Adding the dataflow engine in order to enable filtering and partial monitoring adds
overheads in terms of area and power. We use McPat \cite{mcpat-micro09} to get
a first-order estimate of these area and power overheads in a 40 nm technology
node. McPat estimates the main core area as 2.71 mm$^2$ and the peak power usage as
965 mW averaged across all benchmarks. The average runtime power usage was 385 
mW. These area and power numbers consist of the core and
L1 cache, but do not include L2 cache, memory controller, and other
peripherals. The power numbers include dynamic as well as static (leakage)
power. For the dataflow engine, we modeled the ALUs used for address
calculation, the dataflow flag register file and cache, the configuration
tables, and the filter decision table. These were modeled using the
corresponding memory and ALU objects in McPat. We
note that this is only a rough area and power estimate since components such as the
wires connecting these modules have not been modeled. However, this gives a
sense of the order-of-magnitude overheads involved with implementing our
approach.

Our results show that an additional 0.197 mm$^2$ of silicon area is needed, an
increase of 7\% of the main core area. Table~\ref{tab:evaluation.area_power}
shows the peak and runtime power overheads averaged across all benchmarks with
a 50\% monitoring overhead target. The peak power is 5-7 mW, which is
less than 1\% of the main core's peak power usage. Similarly, the average runtime power is 3-4 
mW, corresponding to about 1\% of the main core's runtime power.

### 2014-05-31 introduction.tex
Various typos/fixes.

% Section summary
This paper is organized as follows. Section~\ref{sec:monitoring} presents the
parallel run-time monitoring model that this paper assumes.
Section~\ref{sec:optimizations} presents the high-level approach of how we aim
to reduce overheads through and Section~\ref{sec:arch} describes our
hardware architecture that supports this. Section~\ref{sec:policies}
investigates the design space for dropping decisions for partial monitoring.
Section~\ref{sec:evaluation} presents our evaluation methodology and
results. Finally, we discuss related work in Section~\ref{sec:related} and
conclude in Section~\ref{sec:conclusion}.

>>>>

% Section summary
This paper is organized as follows. Section~\ref{sec:monitoring} presents the
parallel run-time monitoring model that this paper assumes.
Section~\ref{sec:optimizations} presents the high-level approach of how we aim
to reduce overheads through filtering and partial monitoring and Section~\ref{sec:arch} describes our
hardware architecture that supports this. Section~\ref{sec:policies}
investigates the design space for dropping decisions for partial monitoring.
Section~\ref{sec:evaluation} presents our evaluation methodology and
results. Finally, we discuss related work in Section~\ref{sec:related} and
conclude in Section~\ref{sec:conclusion}.

### 2014-05-31 optimizations.tex
Various typos/fixes.

Although the monitoring core operates in parallel to the main core, the
overheads of programmable monitoring can be high due to the multi-cycle
operation of the monitoring core.
 In order to reduce these overheads, we present two
optimizations. The first optimization is to filter out monitoring events which relate to
null metadata which do not need to processed by the monitoring core (Section~\ref{sec:optimizations.filter}). Second, in order to allow
further reduction in overheads, we propose to perform partial monitoring. This
allows us to further reduce the overheads of monitoring by trading-off
monitoring coverage. Figure~\ref{fig:optimizations.overview} shows an overview
of
these ideas.

@@

Dropping monitoring operations implies that some functionality of the
monitoring scheme has been lost.  This may cause either false negatives, where
an error that occurs in the main program's execution is not detected, or false
positives, where the monitor incorrectly believes an error has occurred.  For
example, a false positive can occur for array bounds check if an event is
dropped that handles copying a pointer. Since the new pointer does not have
correct base and bound address metadata, it will incorrectly cause errors to be
raised.  We accept false negatives as the
loss in coverage that we trade off in order to limit overheads.  However, we
must safely drop monitoring events in such a way as to avoid false positives so
that the system does not incorrectly raise an error.

>>>>

Although the monitoring core operates in parallel to the main core, the
overheads of programmable monitoring can be high due to the multi-cycle
operation of the monitoring core.
 In order to reduce these overheads, we present two
optimizations. The first optimization is to filter out monitoring events which relate to
null metadata which do not need to be processed by the monitoring core (Section~\ref{sec:optimizations.filter}). Second, in order to allow
further reduction in overheads, we propose to perform partial monitoring. This
allows us to further reduce the overheads of monitoring by trading-off
monitoring coverage. Figure~\ref{fig:optimizations.overview} shows an overview
of
these ideas.

@@

Dropping monitoring operations implies that some functionality of the
monitoring scheme has been lost.  This may cause either false negatives, where
an error that occurs in the main program's execution is not detected, or false
positives, where the monitor incorrectly believes an error has occurred.  For
example, a false positive can occur for array bounds check if an event is
dropped that handles copying a pointer. Since the new pointer does not have
correct base and bound address metadata, it will incorrectly cause errors to be
raised.  We accept false negatives as the
loss in coverage that we trade-off in order to limit overheads.  However, we
must safely drop monitoring events in such a way as to avoid false positives so
that the system does not incorrectly raise an error.

### 2014-05-31 policies.tex
Various typos/fixes.

Our goal with enabling partial monitoring is to allow for a specified overhead target or budget to be met while maximizing the amount of monitoring performed.
This is done by dynamically dropping a portion of the monitoring operations. In this section, we discuss in detail how this dropping decision is
made. Specifically, we first discuss how to measure the run-time monitoring
overheads and when monitoring events need to be dropped in
Section~\ref{sec:policies.slack}. Next, we investigate trade-offs created by
selecting which events are dropped. Different policies in determining which
events can be dropped create a trade-off between how closely the overhead
budget is met and the monitoring coverage achieved.

@@

Figure~\ref{fig:policies.stm} shows a hardware slack tracking module for
keeping track of slack. The slack tracking module uses a counter that increments on every $k$-th
cycle of the main core. This $k$ can be calculated by taking the reciprocal of
the target budget. For example, if the target budget is 20\%, then the counter
increments on every 5th clock cycle. The value of this counter is the
accumulated slack. Whenever the main core is stalled due to the monitor, slack
is decremented. This can be difficult to determine precisely due to difficult
to measure overheads such as contention for shared memory. However, we have
found that using only the stalls due to FIFO back pressure works well in
practice.

@@

One method to eliminate this wasted work is to only make dropping decisions at
the root of these metadata flows. That is, we will decide to either monitor or
not monitor an entire metadata flow. We refer to this dropping decision policy
as \emph{source-only dropping} and we refer to the previous policy of 
dropping any event as \emph{unrestricted dropping}.
Figures~\ref{fig:policies.all_drop} and \ref{fig:policies.source_drop} show a
comparison of where dropping decisions are made for these two policies. Source
dropping will
result in no wasted work and thus better coverage. However, because of the coarser-grained decision, it
may be more difficult to closely match overheads. Source dropping is expected
to work well when there are a large number of independent metadata flows.

>>>>

Our goal with enabling partial monitoring is to allow for a specified overhead target or budget to be met while maximizing the amount of monitoring performed.
This is done by dynamically dropping a portion of the monitoring operations. In this section, we discuss in detail how this dropping decision is
made. Specifically, we first discuss how to measure the run-time monitoring
overheads and when monitoring events need to be dropped in
Section~\ref{sec:policies.slack}. Next, we investigate trade-offs created by
selecting which events are dropped (Section~\ref{sec:policies.events}). Different policies in determining which
events can be dropped create a trade-off between how closely the overhead
budget is met and the monitoring coverage achieved.

@@

Figure~\ref{fig:policies.stm} shows a hardware slack tracking module for
keeping track of slack. The slack tracking module uses a counter that increments on every $k$-th
cycle of the main core. This $k$ can be calculated by taking the reciprocal of
the target budget. For example, if the target budget is 20\%, then the counter
increments on every 5th clock cycle. The value of this counter is the
accumulated slack. Whenever the main core is stalled due to the monitor, slack
is decremented. This can be difficult to determine precisely due to the
difficulty in measuring certain overheads such as contention for shared memory.
However, we have found that using only the stalls due to FIFO back pressure
works well in practice.

@@

An alternative dropping policy which eliminates this wasted work is to only make dropping decisions at
the root of these metadata flows. That is, we will decide to either monitor or
not monitor an entire metadata flow. We refer to this dropping decision policy
as \emph{source-only dropping} and we refer to the previous policy of 
dropping any event as \emph{unrestricted dropping}.
Figures~\ref{fig:policies.all_drop} and \ref{fig:policies.source_drop} show a
comparison of where dropping decisions are made for these two policies. Source
dropping will
result in no wasted work and thus better coverage. However, because of the coarser-grained decision, it
may be more difficult to closely match overheads. We expect source dropping 
to work well when there are a large number of independent metadata flows.
